---
title: "Practice with ITT"
output: html_notebook
---

```{r load packages}
library(data.table)
```

```{r load data}
d <- fread("http://hdl.handle.net/10079/70rxwqn")
```

This is the data that is described in Green and Gerbers' *Field Experiments* on page 158. 

```{r}
## Select one-person households that were either pure controls or canvass only
## this just keeps the data that we want to have. 
d <- d[onetreat == 1 & mailings == 0 & phongotv == 0 & persons == 1, ]

## Let's make the var names consistent with the book
setnames(d, "v98", "VOTED")
setnames(d, "persngrp", "ASSIGNED")
setnames(d, "cntany", "TREATED")

## Keep only the data that Green and Gerber Use

d <- d[ , .(VOTED, ASSIGNED, TREATED)]
d
```

# Estimate the ITT 

We can do this two ways: Using means and using models. 


## First lets do it using means:
Can you make back the table 5.2 from Green and Gerber using only the mean function? Here, you should just use group-by calls, and means on the outcome to produce the ITT. Don't worry about the standard errors. 

1. Estimate the ITT using means

```{r}
ITT <- d[, mean(VOTED), ASSIGNED]
ITT
  
```

2. What proportion of those ASSIGNED were actually treated? 

```{r}
ITT_d <- d[ASSIGNED==1, mean(VOTED), TREATED]
ITT_d
```

Can you use these two quantities to produce a CACE? 

```{r}
CACE <- ITT[ASSIGNED==1 , V1, ] / ITT_d[TREATED==1, V1]
CACE
```

## Second, let's do it using models

1. Estiamte the ITT using a `lm` call. 
```{r}
ITT <- d[, lm(VOTED~ASSIGNED)]
ITT
  
```
2. Estiamte the ITT_d using a `lm` call. 
```{r}
ITT_d <- d[, lm(VOTED~TREATED)]
ITT_d
  
```
3. Divide these coefficients. 

```{r}

(ITT$coefficients[1] + ITT$coefficients[2]) / (ITT_d$coefficients[1] + ITT_d$coefficients[2])

```